### 2020-01-30 introduction.tex
added methods section
mwm initial pass

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training and/or testing error.
There are many problems with this approach.
Well-known problems with just examining training/testing curves include that 
they give very limited insight into the overall properties of the model, 
they do not take into account the (often extremely large human and CPU/GPU) time for hyperparameter fiddling,
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
%XXX SOMETHING ELSE, 
and so on.
A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pre-trained model}---and be told to use it.

Na\"{\i}vely---but in our experience commonly, among both ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
This may be true in worst-case theory, but ML models are used in practice, and there is a need for theory to guide that practice.
Moreover, if ML is to become an industrial process, then the process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what the loss of values of the hyperparameters were, how precisely the model was regularized, etc.
%
Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pre-trained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.

In this paper, we present and apply techniques to evaluate the generalization properties of large-scale state-of-the-art pre-trained neural network (NN) models.%
\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pre-trained NN models.}
To do so, we consider a large suite of publicly-available models from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pre-trained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
XXX.  MORE DETAILS.
Importantly, for all of these models, we have no access to training data or testing data.

XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.

In more detail, our main contributions are the following.
\begin{itemize}
\item XXX TECHNICAL THING 1
\item XXX TECHNICAL THING 2
\item XXX TECHNICAL THING 3
\end{itemize}
Perhaps our most improtant contribution, however, is just doing things different, etc. ...

>>>>

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training and/or testing error.
There are many problems with this approach.
Well-known problems with just examining training/testing curves include that 
they give very limited insight into the overall properties of the model, 
they do not take into account the (often extremely large human and CPU/GPU) time for hyperparameter fiddling,
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
and so on.
A somewhat less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may access to neither the training data nor the testing data.
Instead, one may simply be given a model that has already been trained---we will call such an already-trained model a \emph{pretrained model}---and be told to use it.

Na\"{\i}vely---but in our experience commonly, among both ML practitioners and ML theorists---if one does not have access to training or testing data, then one can say absolutely nothing about the quality of a ML model.
This may be true in worst-case theory, but ML models are used in practice, and there is a need for a \emph{practical theory} to guide that practice.
Moreover, if ML is to become an industrial process, then that process will become siloed: some groups will gather data, other groups will develop models, and still other groups will use those models.
The users of models can not be expected to know the precise details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
%
Having metrics to evaluate the quality of a ML model in the absence of training and testing data and without any detailed knowledge of the training and testing process---indeed, having theory for pretrained models, to predict how, when, and why such models can be expected to perform well or poorly---is clearly of interest.

In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.%
\footnote{We reiterate: One could use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
\michael{XXX.  MORE DETAILS.}
Importantly, for all of these models, we have no access to training data or testing data.

\michael{XXX.  LIST PLACES WHERE THEY ARE AVAIALBLE, HERE OR IN NEXT SECTION.}

XXX.  
To do this, we will compute a variety of \emph{quality metrics} based on the spectral properties of the layer weight matrics.
Note that unlike traditional ML approaches, however, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), and \emph{we do not aim to evaluate a single model} (e.g., by training with differing hyperparmeters). 
Instead, we want to examine different models a common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.

In more detail, our main contributions are the following.
\begin{itemize}
\item XXX TECHNICAL THING 1
\item XXX TECHNICAL THING 2
\item XXX TECHNICAL THING 3
\item XXX TECHNICAL THING 4
\end{itemize}

\paragraph{Organization of this paper.}

We start in Section~\ref{sxn:background} and Section~\ref{sxn:methods} with background and an overview of our general approach.
In Section \ref{sxn:cv}, we study three well-known widely-available DNN CV architectures (the VGG, ResNet, and DenseNet series of models); and we provide an illustration of our basic methodology, both to evaluate the different metrics against reported test accuracies and to use quality metrics to understand model properties.
Then, in Section~\ref{sxn:nlp}, we look at several variations of a popular NLP DNN architecture (the OpenAI GPT and GPT2 models); and we show how model quality and properties vary between several variants of GPT and GPT2, including how metrics behave similarly and differently.
Then, in Section \ref{sxn:all_cv_models}, we present results based on an analysis of hundreds of pretraind DNN CV models, showing how well each metric is correlated with the reported test accuracies, and how the Alpha-Norm metric(s) perform remarkably well.
Finally, in Section~\ref{sxn:conc}, we provide a brief discussion and conclusion.

### 2020-01-30 methods.tex
mwm initial pass

We assume we are given several pretrained Deep Neural Networks (DNNs), as part of a similar architecture.
We would like to estimate the trends in the reported test / generalization accuracy accross a series of similar archtectures.  
and without Batch Normalization, trained on ImageNet, and widely available in the pyTorch distribution.

\paragraph{Empirical Complexity Metrics}
To do this, we will compute a variety of \emph{Complexity Metrics} based on the Product Norm of the layer weight matrics.
Note that unlike traditional ML approaches, however, we do not seek a bound on the complexity (i.e. test error), 
nor are we trying to evaluating a single model with differing hyperparmeters.  We wish to examine different models a 
common architecture series. And, also, compare different architectures themselves.

Let us write the Energy Landscape (or optimization function) for a typical DNN with $L$ layers as
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
\label{eqn:dnn_energy}
\end{equation}

with activation functions $h_{l}(\cdot)$,  weight matrices $\mathbf{W}_{l}$, and the biases $\mathbf{b}_{l}$.

The model has been (or will be) trained on (unspecified) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, 
using Backprop, by minimizing some (also unspecified) loss function $\mathcal{L}()$.  Moreover, we expect that most well trained,. production quality models will employ 1 or more forms of on regularization, such as Batch Normalization, Dropout, etc, and will also contain additional structure such as Skip Connections etc. Here, we ignore these details, and focus only on the weight matrices.

Each layer contains by one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or the 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers.  (For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity

\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
 $\Vert\mathbf{W}\Vert$ is a matrix norm.   We will actually compute the log Complexity, which takes the form 
of an Average Log Norm:
\begin{eqnarray*}
\log\mathcal{C} &\sim& \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert
\end{eqnarray*}

Here, we will consider the following Norms:

\begin{itemize}
 \item Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j}$
 \item Spectral Norm:  $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
 \item $\alpha-$Norm (or Shatten Norm) $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda^{\alpha}$,
\end{itemize}

where $\lambda_{i}$ is the $i-th$ eigenvalue of the \emph{Empirical Correlation Matrix}

\begin{equation}
\mathbf{X}=\mathbf{W}^{T}\mathbf{W}
\end{equation}

and $\lambda_{max}$ is the maximum eigenvalue. These eigenvalues are the square of the singular values $\sigma_{i}$ of $\mathbf{W}$ :  $\lambda_{i}=\sigma^{2}_{i}$.

The exponent $\alpha$ is the power law exponent that arises in our \emph{Theory of Heavy Tailed Self Regularization}, and is the determined by fitting the Empirical Spectral Density (ESD) of $\mathbf{X}$--i.e. a histogram of the eigenvalues--$\rho(\lambda)$ to a truncated power law

\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}
\end{equation}

We will also consider an approximate capacity metric, $\hat{\alpha}$, shown previously to correlate well with the trends in reported test accuracies of pretrained DNNs \cite{MM}

\begin{itemize}
 \item $\hat{\alpha}=\alpha\log\lambda_{max}\approx\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$
\end{itemize}

which approximates the log $\alpha-$Norm for both Very Heavy Tailed weight matrices ($\alpha < 2)$  and reasoably well for finite size, Moderately Heavy Tailed ones $\alpha\in[2,5]$.

\paragraph{Spectral Analysis of Convolutional 2D Layers}
While we can easily analyze Linear layers, there is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  A Conv2D layer is a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel), and $in\_ch,out\_ch$ input and output channels, respectively.  Typically, the $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each.  Usually $in\_ch\le out\_ch$.

There are at least 3 different approaches to computing the Singular Values Decomposition(s) (SVD) of an Conv2D layer
\begin{enumerate}
\item run SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values. 
\item stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values
\item Compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients\cite{Long2019}.  This leads to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
\end{enumerate}

Each method has tradeoffs.  While, in principle, 3. is mathematically sound, it is computationally expensive. 
For this study, because we are computing tens of thousands of calculations, we select $1.$, which is 
numerically the fastest and easiest to reproduce.\footnote{We will provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with option 3 for the SVD of the Conv2D.}

To verify that our approach is meaningful, we need to see the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data.  We examine typical Conv2D layer for the pretained AlexNet model (distributed with pyTorch).  Figure \ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.  The red line displays the best fit to a random matrix, using the Marchenko pastur theory\cite{MM}.  We can see the random matrix model does not describe the ESD very well. For comparison, Figure \ref{fig:alexnet2} shows the ESD of the same matrix, randomly shuffled; here looks similar to the red line plot of the orginal ESD.  In fact, the empircal ESD is better modeled with a truncated power law distribtion.

Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--
only the correlation matrix $\mathbf{X}$ is. If $\mathbf{W}$ was, 
then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original
nor is it heavy tailed.  So we can rule out $\mathbf{W}$ being heavy tailed.

\begin{figure}[H]
   \centering
   \subfigure[Actual ESD]{
     \includegraphics[scale=0.5]{img/alexnet1.png} 
     \label{fig:alexnet1}
   }
   \subfigure[ESD of randomly shuffled matrix]{
      \includegraphics[scale=0.5]{img/alexnet2.png}
      \label{fig:alexnet2}
   }
   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized}
   \label{fig:alexnet}
\end{figure}

These plots tell us that the pre-activation maps of the Conv2D contains significant correlations learned from the data.  
By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
the smaller the exponent $\alpha$, the more correlation in the weight matrix.

\paragraph{Normalization of Empirical Matrices}.  
The normalization does not affect the Power Law fits because the Heavy Tailed exponent $\alpha$ is scale-invariant.
As are metrics like the Stable Rank and our MP Soft Rank. In this study, however, all of the metrics strongly depend 
on the scale of the weight matrix.
Formally, to apply RMT, we would usually define Correlation Matrix with $\dfrac{1}{N}$ normalization,
and assume that either the variance of $\mathbf{X}$ is unity, $\sigma^{2}=1$, or is known.
\footnote{And for formal proofs of the Heavy Tailed results, we need a different normalization,$\dfrac{1}{N^{1-\alpha}}$}
But for these empirical studies, the pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$ ,
with the variance already normalized to $\sqrt{\frac{1}{N}}$, or some variant of this such as Glorot (or Xavier) Normalization\cite{GloRot},
or as $\sqrt{\frac{2}{N*k^2}}$ for Convolutional 2D Layers. 
We do not normalize (or renormalize) the Empirical Correlation Matrices, and use them as-is,
except that we rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $\frac{k}{\sqrt{2}}$ so that they
are on the same scale as the Linear layers.

Note that in some pretrained models, like the pyTorch VGG models, the Linear weight matrices appear to have been 
initialized with variance $\sigma\sim0.01$.  We do not attempt to correct for this, and, instead, simply try to treat all
models the same.

COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else

COMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.

\paragraph{The WeightWatcher Tool}

\paragraph{Organization of this paper}

We want to study the effective of Product Norm Metrics in describing trends in the
generalization accuracyof real-word, production quality, pretrained Deep Neural Networks.
In Section \ref{sxn:cv}, we study three well known, and widely available DNN CV architectures:
the VGG series, the ResNet series, and the DenseNet series of models.  We lay out the
basic methodology we use to evaluate the different metrics against the reported test accuracies. 
We also introduce the notion of Correlation Flow, compare the three pretrained architectures in detail.  
In Section  \ref{sxn:nlp}, we then look at several variations of a popular NLP DNN architecture,
the OpenAI GPT and GPT2 models.  We discuss when and how the different metrics apply more generally.
Finally, in Section \ref{sxn:all}, we examine over 100 pretraind DNN CV models, and show
how well each metric is correlated with the reported test accuracies across different models
in an architeture series.  
----

>>>>

Let us write the Energy Landscape (or optimization function, parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
\label{eqn:dnn_energy}
\end{equation}
We assume we are given several pretrained Deep Neural Networks (DNNs), e.g., as part of an architecture series.
The models have been trained on (unspecified and unavailable) labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing some (also unspecified and unavailable) loss function $\mathcal{L}()$. 
We expect that most well-trained, production-quality models will employ one or more forms of on regularization, such as Batch Normalization, Dropout, etc., and many will also contain additional structure such as Skip Connections, etc. 
Here, we will ignore these details, and will focus only on the weight matrices.

\paragraph{DNN Quality Metrics.}

Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{L}$, and/or 2D feature maps $\mathbf{W}_{i,L}$ extracted from 2D Convolutional layers. 
(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below.) 
%
\michael{
XXX.  PROBABLY WE DO NOT WANT THE FOLLOWING.  INSTEAD, JUST PRESENT METRICS AND POINT TO PREVIOUS PAPERS FOR JUSTIFICATION.  E.G., WE DONT WANT TO JUSTIFY THE IMPLICIT STATISTICAL INDEPENDENCE ASSUMPTION HERE.  IF OKAY, JUST DELETE THIS RED BLOCK.
We assume the layer weight matrices are statistically independent, allowing us to estimate the Complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm, which resembles a data dependent VC complexity
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert\cdots\Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\mathbf{W}$ is an $(N\times M)$ weight matrix, with $N\ge M$, and 
 $\Vert\mathbf{W}\Vert$ is a matrix norm.   We will actually compute the log Complexity, which takes the form 
of an Average Log Norm:
\begin{eqnarray*}
\log\mathcal{C} &\sim& \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert\cdots\log\Vert\mathbf{W}_{L}\Vert
\end{eqnarray*}
}
%
We have examined a large number of possible quality metrics.
The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about }
We consider the following metrics.
\begin{itemize}
\item 
Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{W}\Vert^{2}_{2}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
\michael{Do we use norm or log norm, here and with other norms.  Worth being very explicit and very consistent.}
\item 
Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
\michael{We should probably call the spectral norm $\Vert\mathbf{W}\Vert_{2}$, or change Frob norm notation and have an explicit remork.  Let's decide.}
\item 
Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
\item 
$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$,
\michael{Be clear about how we average this across layers.}
\end{itemize}
The first two metrics are well-known in ML.
The last two deserve special mention.
For all these metrics, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
$ %% $$
\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
$ %% $$
and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
For the last two metrics, the exponent $\alpha$ is the power law exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool (\cite{weightwatcher_package}) to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated power law
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  .
\end{equation}
\michael{We need to be clear that this is a truncated power law fit and that $\lambda_{max}$ comes from that and not the largest empirical eigenvalue.}
The Weighted Alpha Metric was introduced previously~\cite{MM20_SDM}, where (on a much smaller set of data than we consider here) it was shown to correlate well with the trends in reported test accuracies of pretrained DNNs.
Based on this, here we introduce and evaluate the $\alpha$-Norm metric.
One would expect $\hat{\alpha}$ to approximate the log $\alpha$-Norm very well for $\alpha < 2$ and reasoably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.

\michael{Need to say that such large $\alpha$ values don't mean much.}

\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}

\paragraph{Spectral Analysis of Convolutional 2D Layers.}

There is some ambiguity in performing spectral analysis on Convolutional 2D (Conv2D) layers.  
A Conv2D layer can be represented as a 4-index tensor of dimension $(w,h,in\_ch,out\_ch)$, specified by an $(w\times h)$ filter (or kernel) and $in\_ch$ / $out\_ch$ input / output channels, respectively (usually $in\_ch\le out\_ch$). 
Typically, $w=h=k$,  giving $(k\times k)$ tensor slices, or \emph{pre-Activation Maps} $\mathbf{W}_{i,L}$ of dimension $(in\_ch\times out\_ch)$ each. 
%
There are at least three different approaches that have been advocated for applying the Singular Values Decomposition (SVD) to an Conv2D layer:
run an SVD on each of the pre-Activation Maps $\mathbf{W}_{i,L}$, yielding $(k\times k)$ sets of $M$ singular values; 
stack the feature maps into a single rectangular matrix of, say, dimension $((k\times k\times out\_ch)\times in\_ch)$, yielding $in\_ch$ singular values;
compute the 2D Fourier Transform (FFT) for each of the $(in\_ch, out\_ch)$ pairs, and run SVD on the resulting Fourier coeffients~\cite{Long2019}, leading to $\sim(k\times in\_ch\times out\_ch)$ non-zero singular values.
Each method has tradeoffs.  
In principle, the third method is mathematically sound, but it is computationally expensive. 
For our analysis, because we are performing tens of thousands of calculations, we select the first method, which is numerically the fastest and is easiest to reproduce.%
\footnote{We provide a Google Colab notebook where all results can be reproduced, with the option to redo the calculations with the third option for the SVD of the Conv2D.}

\paragraph{XXX.}
XXX.  THIS VERIFICATION IS NOT ABOUT CONV LAYERS, IT IS ABOUT PL MORE GENERALLY, CORRECT?  WE SHOULD CLARIFY AND SQUISH.
To verify that our approach is meaningful, we need to confirm that the ESD is neither due to a random matrix, nor due to unsually large matrix elements, but, in fact, captures correlations learned from the data. 
We examine typical layer for the pretained AlexNet model (distributed with pyTorch). 
Figure~\ref{fig:alexnet1} displays the ESD for the first slice (or matrix $\mathbf{W}$) of the third Conv2D layer, extracted from a 4-index Tensor of shape $(384, 192, 3, 3)$.  The red line displays the best fit to a random matrix, using the Marchenko pastur theory~\cite{MM}.  We can see the random matrix model does not describe the ESD very well. For comparison, Figure \ref{fig:alexnet2} shows the ESD of the same matrix, randomly shuffled; here looks similar to the red line plot of the orginal ESD.  In fact, the empircal ESD is better modeled with a truncated power law distribtion.
\michael{We may want to give a one-sentence summary of this par and fig at the end of the previous par.}

\begin{figure}[H]
   \centering
   \subfigure[Actual ESD]{
     %\includegraphics[scale=0.5]{img/alexnet1.png} 
     \includegraphics[scale=0.25]{img/alexnet1.png} 
     \label{fig:alexnet1}
   }
   \subfigure[ESD of randomly shuffled matrix]{
      %\includegraphics[scale=0.5]{img/alexnet2.png}
      \includegraphics[scale=0.25]{img/alexnet2.png}
      \label{fig:alexnet2}
   }
   \caption{ESD of AlexNet Conv2D pre-Activation map for Layer 3 Slice 1, actual and randomized.
            \michael{Need better figs here.}
           }
   \label{fig:alexnet}
\end{figure}

Although the ESD is \emph{Heavy Tailed}, this does not imply that the orginal matrix $\mathbf{W}$ is itself heavy tailed--only the correlation matrix $\mathbf{X}$ is. 
If $\mathbf{W}$ was, then it would contain 1 or more unusually large matrix elements, and they would dominate the ESD.  
Of course the randomized $\mathbf{W}$ would also be heavy tailed, but its ESD neither resembles the original nor is it heavy tailed. 
So we can rule out $\mathbf{W}$ being heavy tailed.
\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}

These plots tell us that the pre-activation maps of the Conv2D contains significant correlations learned from the data.  
By modeling the ESD with a power law distribution $\lambda^{\alpha}$, we can characterize the amount of correlation learned;
the smaller the exponent $\alpha$, the more correlation in the weight matrix. 
\michael{These comments seem out of place, since they hold more generally than for the Conv2D layers.}

\paragraph{Normalization of Empirical Matrices.}  
Normalization is an important, if underappreciated, practical issue.
Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
\footnote{For formal proofs of Heavy Tailed results, one typically needs a different normalization such as $1/N^{1-\alpha}$.}
Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.

\michael{MOVE TO LATER: COMMENT ON HOW LOG NORM first and last layers behave, maybe somewhere else.}

\michael{CMOVE TO LATER: OMMENT ON HOW LOG PORM  for GPT includes unusually high alpha, not meaningful other than to show the trend.}

\paragraph{The WeightWatcher Tool.}

We compute the metrics using the WeightWatcher tool (version 0.2.7), and we provide Jupyter notebooks in the github repo for this paper~\cite{repo}, making the results fully reproducible.

\michael{XXX.  FEW SENTENCES ABOUT REPRODUCIBILITY MORE GENERALLY HERE.}
